---
title: "Project on Dataset Analysis"
author: "Subash Nepal"
date: "2024-07-24"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
# INTRODUCTION
This project describes about the comprehensive analysis of the Student Performance Dataset.

# SOURCE
This dataset was extracted from **Kaggle.com**. 
**https://www.kaggle.com/datasets/rabieelkharoua/students-performance-dataset**

# About Dataset 
This dataset contains comprehensive information on 2,392 high school students. This dataset offers a comprehensive view of the factors influencing students' academic performance, making it ideal for educational research, development of predictive models, and statistical analysis.

# Table of COntents 
**1. Student Information**
  
  >. Student ID
  
  >. Demographic Details
    
  >. Study Habits
   
**2. Parental Involvement**

**3. Extracurricular Activities**

**4. Academic Performance**

**5. Target Variable: Grade Class**

# 1. Student Information
  **> Student ID (1001 to 3392)**
  
  **> Demographic Details**
     
     *Age(From 15 to 18)*
     
     *Gender(0-Male and 1-Female)*
     
     *Ethnicity(0-Caucasian, 1-African American, 2-Asian and 3-Other)*
   
     *Parental Education(0-None, 1-High School, 3-Bachelors and 4-Higher)*
     
  **> Study Habits**
  
     *Study Time Weekely (Range from 0-20)*
     
     *Absences/Year (Ranging from 0 to 30)*
     
     *Tutoring Status (0-No and 1-Yes)*
     
# 2. Parental Involvement
  **> Parental Support**
      
      0 - None
      
      1 - Low
      
      2 - Moderate
      
      3 - High
      
      4 - Very High

# 3. Extracurricular Activities
  **> Extracurricular**
  
    Participation in ECA (0-NO and 1-YES)
  
  **> Sports**
  
    Participation in Sports (0-NO and 1-YES)
  
  **> Music**
  
    Participation in Music Activities (0-NO and 1-YES)
  
  **> Volunteering**
  
    Participation in Volunteering (0-NO and 1-YES)
    
# 4. Academic Performance
  **> GPA**
    
    Grade Point Average on a scale from 2.0 to 4.0, influenced by study habits, parental involvement and ECA.
    
# 5. Target Variable: Grade Class
  **> Grade Class: Classification of students' grades based on GPA**
  
    . 0 - 'A' (GPA >= 3.5)
    
    . 1 - 'B' (3.0 <= GPA < 3.5)
    
    . 1 - 'C' (2.5 <= GPA < 3.0)
    
    . 3 - 'D' (2.0 <= GPA < 2.5)
    
    . 4 - 'F' (GPA < 2.0)
    
  
# Dataset Loading
```{r}
file = read.csv("Student.csv")
```
# First Five Values of Dataset
```{r}
head(file)
```
# Structure of Column of Dataset
```{r}
str(file)
```
# Cleaning of data
```{r}
sum(is.na(file))

sum(is.null(file))
```
**Here the *N/A* or *Null* value is no where to be found in the dataset. Hence, we can conclude that the dataset is almost clean.**

# Summary of Dataset
```{r}
summary(file)
```

# Exclusion of Unnecessary Column 
Here *Tutoring* and *ECA section(Sports, Music and Volunteering)* is not required. So these need to extract these columns from the dataset.
```{r}
file = subset(file, select = -c(Tutoring, Extracurricular, Sports, Music, Volunteering))
View(file)
head(file)
```

# Shape of Data
```{r}
dim(file)
```

# Number of Student at Different Age
```{r}
table(file$Age)
```
Students of different ages:

  15 yrs old = 630
  
  16 yrs old = 593
  
  17 yrs old = 587
  
  18 yrs old = 582

# Number of Males and Females
```{r}
table(file$Gender)
```
Here the number of Male student is **1170** and Female student is **1222**.

# Number of Males and Females at Different Ages
```{r}
attach(file)
M=file$Age[Gender==0]
M15= M[M==15]
length(M15)
```
There are 338 Male Students of age 15.

```{r}
M=file$Age[Gender==0]
M16= M[M==16]
length(M16)
```
There are 283 Male student of age 16.

```{r}
M=file$Age[Gender==0]
M17= M[M==17]
length(M17)
```
There are 272 Male student of age 17.
```{r}
M=file$Age[Gender==0]
M18= M[M==18]
length(M18)
```
There are 277 Male student of age 18.

```{r}
F=file$Age[Gender==1]
F15= F[F==15]
length(F15)
```
There are 292 Female student of age 15.

```{r}
F=file$Age[Gender==1]
F16= F[F==16]
length(F16)
```
There are 310 Female student of age 16.

```{r}
F=file$Age[Gender==1]
F17= F[F==17]
length(F17)
```
There are 315 Female student of age 17.
```{r}
F=file$Age[Gender==1]
F18= F[F==18]
length(F18)
```
There are 305 Female student of age 18.

# Plot of Student on basis of age and gender
```{r echo=FALSE}
library(ggplot2)
data = data.frame(
  Gender = rep(c("Male", "Female"), each = 4),
  Age = rep(c(15, 16, 17, 18), times = 2),
  Count = c(338, 283, 272, 277, 292, 310, 315, 305)
)
ggplot(data, aes(x = Age, y = Count, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Age Distribution by Gender",
       x = "Age",
       y = "No. of Students") +
  theme_minimal()+
  scale_fill_manual(values = c("Male" = "maroon", "Female" = "skyblue")
)
```

This bar graph demonstrates the number of students of a age on the basis of the gender.

```{r}

barplot(file$GPA,file$StudyTimeWeekly,
        main = "Affect of Study Time in GPA",
        ylab = "GPA",
        
        xlab = "Study Time")
```





```{r}
file = subset(file, select = -c(Ethnicity))
```

```{r}
model1 <- glm(file$GPA~file$StudyTimeWeekly+file$Absences+file$ParentalSupport)
summary(model)
```